---
title: "Sam Is Learning R"
author: "Sam Church"
description: "Final Study Review"
output:
  distill::distill_article:
    self_contained: false
---
## Setup

for the purposes of examples:

```{r}
library(tidyverse)
```

fake data creation:

```{r}
data_here <- read_csv("data/union.csv")

data_here <- data_here |>
  mutate(
    independent_var = if_else(union_now == 1, "Treated", "Control"),
    dependent_var = rr
  )
```


## Bootstrap

  Bootstraps are easy to generate!

  A **bootstrap** mimics large-scale sampling by resampling many times from the sample itself. It is done WITH replacement,     otherwise it would just mimic the sample and nothing would differ.

  1 make sure you've loaded the infer package
```{r}
library(infer)
```

  2 use specify to clarify the dependent and independent varaibles:
```{r}
bootstrap <- data_here |>
  specify(dependent_var ~ independent_var)
```
  
  3 now we can actually do the work to generate the bootstrap:
```{r}
bootstrap <- data_here |>
  specify(dependent_var ~ independent_var) |>
  generate(reps = 1000, type = "bootstrap")
```

  4 now that we've done the work to actually generate the bootstrap, we can do differnet things with it.

  A. For example, if we're looking for the ATE, we can calculate that:
```{r}
bootstrap <- data_here |>
  specify(dependent_var ~ independent_var) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means", c("Treated", "Control"))
```

  B. We can also just calculate the mean:

```{r}
bootstrap <- data_here |>
  specify(response = dependent_var) |>
  ##when doing calculations involving just ONE VARIABLE, change the specify plank to clarify what variable you're referencing!
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean")
```









