---
title: "Sam Is Learning R"
author: "Sam Church"
description: "Final Study Review"
output:
  distill::distill_article:
    self_contained: false
---
## Setup

for the purposes of examples:

```{r}
library(tidyverse)
```

fake data creation:

```{r}
data_here <- read_csv("data/union.csv")

data_here <- data_here |>
  mutate(
    independent_var = if_else(union_now == 1, "Treated", "Control"),
    dependent_var = rr
  )
```


## Bootstrap

Bootstraps are easy to generate!

A **bootstrap** mimics large-scale sampling by resampling many times from the sample itself. It is done WITH replacement,     otherwise it would just mimic the sample and nothing would differ.

**First** make sure you've loaded infer!
```{r}
library(infer)
```

**Second** use specify to describe the relationship between what you're testing.
```{r}
bootstrap <- data_here |>
  specify(dependent_var ~ independent_var)
```
  
**Third** we can actually generate the bootstrap.
```{r}
bootstrap <- data_here |>
  specify(dependent_var ~ independent_var) |>
  generate(reps = 1000, type = "bootstrap")
```

**Fourth** we can do a lot of different calculations with it!

For example, if we're looking for the ATE, we can calculate that.
```{r}
bootstrap <- data_here |>
  specify(dependent_var ~ independent_var) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means", c("Treated", "Control"))
```

We can also just calculate the mean.

```{r}
bootstrap <- data_here |>
  specify(response = dependent_var) |>
  ##when doing calculations involving just ONE VARIABLE, change the specify plank to clarify what variable you're referencing!
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean")
```









